Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-10 arxiv/diff_cov.tex

\section{The Different Covariates Case}\label{sec_diff}

@@

	\noindent (i) The variance equation \eqref{eq_variance_2task} satisfies the following limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
	\noindent (ii) The bias equation \eqref{eq_bias_2task} satisfies the following limit for any fixed vector $w\in\real^p$ with unit norm,
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following self-consistent equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_1 a_4 - a_2 a_3)}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
Our result extends a well-known result on the inverse of a sample covariance matrix.
In particular, when $X_1$ is zero, we obtain Fact \ref{lem_minv}(ii) as a corollary.
To see this, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to the variance limit.
One can see that both the variance limit and the bias limit depend heavily on both tasks' samples sizes and the covariate shift matrix $M$.
Next, we illustrate how varying sample sizes and covariate shifts impact hard parameter sharing, respectively.

\begin{example}[Sample ratio]\label{ex_sample_ratio}
	We first consider the impact of varying sample sizes.
	Consider the random-effects model from Section \ref{sec_hps}, with both tasks having an isotropic population covariance matrix.

>>>>

\section{Bias and Variance Limits when Tasks have Different Sample Sizes and Covariate Shifts}\label{sec_diff}

@@

	\noindent (i) The variance equation \eqref{eq_variance_2task} satisfies the following limit
			\begin{align}\label{lem_cov_shift_eq}
				\bigabs{\bigtr{\Sigma^{(2)} \bigbrace{(n_1 + n_2){\hat{\Sigma}^{-1}} - (a_1 \Sigma^{(1)} + a_2 \Sigma^{(2)})^{-1}}}} \lesssim p^{-c_{\varphi}}\norm{\Sigma^{(2)}},
			\end{align}
			where $a_1$ and $a_2$ are the solutions of the following self-consistent equations
			\begin{align}
				a_1 + a_2 = 1- \frac{p}{n_1 + n_2},\quad a_1 + \frac1{n_1 + n_2}\cdot \bigbrace{\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2}} = \frac{n_1}{n_1 + n_2}. \label{eq_a12extra}
			\end{align}
	\noindent (ii) The bias equation \eqref{eq_bias_2task} satisfies the following limit for any fixed vector $w\in\real^p$ with unit norm,
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{(n_1+n_2)^2\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - {\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \lesssim p^{-c_{\varphi}},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following self-consistent equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2},
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_1 a_4 - a_2 a_3)}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
Our result extends a well-known result on the inverse of a sample covariance matrix.
In particular, when $X_1$ is zero, we obtain Fact \ref{fact_tr} as a corollary.
To see this, we solve equation \eqref{eq_a12extra} and obtain $a_1 = 0$ and $a_2 = (n_2-p) / n_2$, and apply these solutions to the variance limit.
One can see that both the variance limit and the bias limit depend heavily on both tasks' samples sizes and the covariate shift matrix $M$.
Next, we illustrate how varying sample sizes and covariate shifts impact hard parameter sharing, respectively.

\begin{example}[Sample ratio]\label{ex_sample_ratio}
	We first consider the impact of varying sample sizes.
	Consider the random-effects model from Section \ref{sec_same}, with both tasks having an isotropic population covariance matrix.

### 2020-10-10 arxiv/tools.tex
sec A

\section{Toolbox}\label{app_tool}

@@

\subsection{Random Variables with Bounded Moments}
The following lemma gives (almost) sharp concentration bounds for linear and quadratic forms of bounded supported random variables. Here we recall that the stochastic domination ``$\prec$" was defined in Definition \ref{stoch_domination}.
%Besides the proof of Lemma \ref{prop_entry}, we also use it in several other places of this paper, including the proofs in Section \ref{sec_proof_general} and Section \ref{app_iso_cov}.
%It constitutes the main difference between our proof and the one in \cite{KY2}, where the authors used a large deviation bound for random variables with arbitrarily high moments.

@@

>>>>

\section{Notations and Tools}\label{app_tool}

@@

\subsection{Random Variables with Bounded Moments}

\begin{fact}\label{fact_minv}
	In the setting of Theorem \ref{thm_many_tasks}, with high probability over the randomness of $X$, we have that
	\begin{enumerate}
		\item When $n / p$ converges to $\rho > 1$, the sample covariance matrix $\frac{X^{\top}X}{n}$ is full rank.
		\item The singular values of $Z^{\top}Z$ are greater than $(\sqrt{n} - \sqrt{p})^2 - n \cdot p^{-c_{\varphi}}$ and less than $(\sqrt{n} + \sqrt{p})^2 + n \cdot p^{-c_{\varphi}}$, cf. \citet[Theorem 2.10]{isotropic} and \citet[Lemma 3.12]{DY}.
	\end{enumerate}
\end{fact}

The following lemma gives (almost) sharp concentration bounds for linear and quadratic forms of bounded supported random variables. Here we recall that the stochastic domination ``$\prec$" was defined in Definition \ref{stoch_domination}.
%Besides the proof of Lemma \ref{prop_entry}, we also use it in several other places of this paper, including the proofs in Section \ref{sec_proof_general} and Section \ref{app_iso_cov}.
%It constitutes the main difference between our proof and the one in \cite{KY2}, where the authors used a large deviation bound for random variables with arbitrarily high moments.

@@

 Moreover, we can obtain the following concentration estimates for random noise satisfying \eqref{assmAhigh2}.

 \begin{corollary}\label{cor_calE}
 add corollary

 using Lemma \ref{largedeviation} in Appendix \ref{app_tool} and the fact that all moments of $\varepsilon_i$ exist by Assumption \eqref{assmAhigh2}
 \end{corollary}
